Article Title

### 2014-01-22 plos.tex
corrections

It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component. The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
\section*{Introduction}
\emph{Functional connectivity} is a statistical description of \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.  Functional connectivity is commonly expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons or in the form of joint coactivation patterns.  Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.  To remove the confounding effects of the stimulus, the average stimulus responses are subtracted or otherwise accounted for. Functional connectivity is only a phenomenological description with no simple correspondence to anatomical connectivity, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and how information is encoded and processed at the population level.

Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of the anatomical connectivity \cite{Gerstein:1964}.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.

Neural correlations do not come with ready or unambiguous mechanistic interpretations. They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}. But multineuronal recordings do provide more information than available from an equivalent number of pairs of cells considered separately. For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients. In contrast, the off-diagonal elements of the precision matrix (inverse of the covariance matrix) constitute scaled partial correlations between neuron pairs.  The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.  For this reason, partial correlations have been proposed to more closely reflect the direct effects between components of complex systems and have been used to describe gene interactions \cite{Schafer:2005, Peng:2009} and the interactions between brain regions \cite{Varoquaux:2012, Ryali:2012}.

However, estimation of covariance matrices from large populations presents a number of numerical challenges:  The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically.  This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}. 
The \emph{sample correlation matrix}, which is an unbiased estimate, is often used to estimate the true correlation matrix. It has many free parameters making it highly sensitive to sampling noise. As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.

In statistics, the technique of deliberately imposing a structure on an estimate in order to improve its performance is known as \emph{regularization} \cite{Schafer:2005, Bickel:2006}.    To \sq{impose structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation known as the shrinkage target.  The selection of the optimal target estimate and the optimal amount of shrinkage can be deduced from the training data either analytically or by cross-validation.  Some regularization schemes only perform target selection (\emph{dimensionality reduction, variable selection, or feature selection}) and replace the estimate completely. Others only perform shrinkage toward a single target estimate (\emph{shrinkage estimators}).  Yet others do both target selection and shrinkage.

Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.  All prior empirical findings from neural correlations have relied on sample correlations. Improved estimation is not the only or even the main benefit of regularization.  When we select the best performing estimator among many regularized estimators, we learn something about the system itself:  The structure imposed by the best estimator represents the underlying dependencies in the system more closely than the structures imposed by the other estimators.  We can then examine this structure as a more effective description of the functional connectivity in the circuit than the sample correlation matrix.

@@

To address these questions, in addition to the sample covariance matrix, we evaluated four regularized covariance matrix estimators that imposed different kinds of structure on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage toward a diagonal covariance matrix. 
\item[$C_{\sf factor}$] -- linear shrinkage toward a low-rank matrix produced by factor analysis.
\item[$C_{\sf sparse}$] --  sparse partial correlations, \ie covariance selection.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons combined with partial correlations to several inferred latent units.
\end{description}

First, we demonstrated that, in simulations with known correlation structures, estimators that imposed a matching structure were generally more efficient than other estimators. We therefore reasoned as follows: The most efficient covariance matrix estimator is likely to provide a better picture of the correlation structure than less efficient estimators (See Discussion).

We therefore performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units with partial correlations with the observed cells. We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distance and differences in preferred orientations than noise correlations. In contrast, the inferred negative partial correlations were less selective.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The covariance matrix is defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.  The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.  In the case of noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus.

@@

We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, the targets estimates of these for estimators can correspond to distinct families of graphical models (\figref{02}{\,Row}{row 1}).

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Estimators whose regularization targets represent the structure of the true covariance matrix outperform other estimators.}
        {\bf Row 1.} Graphical representations of the target estimates used by the four respective regularized covariance matrix estimators.  Recorded neurons are denotes as green spheres and latent units as light-shaded spheres.  The edges between them denote non-zero partial correlations.
        In estimator $C_{\sf diag}$, the regularization target is a diagonal matrix.  This target corresponds to a model wherein observed neurons exert no linear effects on one another ({\bf Row 1, A}). 
        In estimator $C_{\sf factor}$, the regularization target is a factor model (low-rank matrix plus a diagonal matrix).  This target describes a model wherein the activity of observed neurons is linearly driven by several latent units ({\bf  Row 1, B}).
        In estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes a model wherein correlations arise due to linear effects exerted by the recorded cells onto one another other ({\bf  Row 1, C}).
        In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of a matrix constructed as the sum of a sparse and a low-rank component. This approximation describes a model wherein correlations arise due to linear affects exerted by the recorded cells onto one another and due to linear effects from several latent units ({\bf  Row 1, D}).
\\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: the diagonal correlation matrix ({\bf  A}), a factor model with two latent units ({\bf B}),  a correlation matrix with 85\%  off-diagonal zeros ({\bf  C}), and a correlation matrix whose inverse is the some a rank-1 (\ie one latent unit) component and a sparse component with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of 1000 observations drawn from simulated random processes with respective correlation matrices from row 2.  The structure of the correlation matrix is much more difficult to discern by eye.
\\
{\bf Row 4:} Regularized correlation matrix estimates computed from the same samples as in Row 3.  For each kind of ground truth, the estimator with the matching structure was used and the regularization parameters optimized by cross-validation using the data in the same sample.   The regularized estimates are closer to truth than sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by each of the five estimators for  various sample sizes. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with the matching structure converged to zero faster than the other estimators when the sample size was large enough to allow the more complex true structures outperform simpler structures.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained  by each of the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons’ variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure of imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).  If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ is expected to get closer to truth than other estimators.

The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
\end{equation}
has two hyperparameters: the number of factors in $L$ and the shrinkage intensity $\lambda$. The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{Row 1, B}).

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by a matrix whose inverse matrix is sparse: 
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros in its off-diagonal.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.  The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).  We will sometimes refer to sparse partial interactions between neurons as simply \sq{interactions}.

>>>>

It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component. The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.
\section*{Introduction}
\emph{Functional connectivity} is a statistical description of \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells.  Functional connectivity is expressed as the degree of correlation, synchronization, or other forms of statistical association between neurons, or in the form of multineuronal coactivation patterns, or by fitting a probabilistic model of the population activity.  Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.  Although functional connectivity is only a phenomenological description with no simple mechanistic interpretation, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and the processing of information at the population level.

Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.  The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.

Neural correlations do not come with ready or unambiguous mechanistic interpretations. They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}. But multineuronal recordings do provide more information than available from an equivalent number of pairs of cells considered separately. For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients. On the other hand, a correlation matrix can be specified through the  partial correlations between pairs of the recorded neurons. The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.  For this reason, partial correlations have been proposed to more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}.

However, estimation of covariance matrices from large populations presents a number of numerical challenges. The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically. This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}. All prior empirical findings from neural correlations have used \emph{sample correlations}.  The \emph{sample correlation matrix} is an unbiased estimate but its many free parameters make it sensitive to the sampling noise. As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.

Estimation can be improved through \emph{regularization},  the technique of deliberately imposing a structure on an estimate in order to reduce its estimation error \cite{Schafer:2005, Bickel:2006}. To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation with fewer free parameters, which is often referred to as the \emph{target estimate}.   The selection of the optimal target estimate and the optimal amount of shrinkage is deduced from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005}  or by cross-validation \cite{Friedman:1989}. An estimator that produces estimates that are, on average, closer to the truth for a given sample size is said to be more \emph{efficient} than other estimators.

Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.

Reduced estimation error is not the only benefit of regularization.  By finding the most efficient estimator for a specific system among many regularized estimators, we learn something about the system itself: Since imposing this structure leads best estimation, the elements of the structure may provide a better representation of the underlying dependencies in the data than previous analyses of sample correlations.

@@

To address these questions, we evaluated four regularized covariance matrix estimators that imposed different structures on the estimate. The estimators are designated as follows:
\begin{description}
\item[$C_{\sf 0}$] -- sample covariance matrix, the unbiased, unregularized estimator.
\item[$C_{\sf diag}$] -- linear shrinkage of correlations toward zero.
\item[$C_{\sf factor}$] -- linear shrinkage of the covariances toward a low-rank matrix, representing inputs from unobserved factors.
\item[$C_{\sf sparse}$] -- sparse partial correlations, \ie a large fraction of the \emph{partial} correlations between pairs of neurons are set to zero.
\item[$C_{\sf sparse+latent}$] -- sparse partial correlations between the recorded neurons \emph{and} linear interactions with a number of latent units.
\end{description}

First, we used simulated data to demonstrate that the selection of the optimal estimator indeed pointed to the true structure of the dependencies in the data.

We therefore performed a cross-validated evaluation to establish which of the four regularized estimators was most efficient for the population activity of dense groups of neurons in mouse primary visual cortex recorded with high-speed 3D random-access two-photon imaging of calcium signals. We found that $C_{\sf sparse+latent}$ consistently outperformed the other estimators.  This estimator revealed a sparse network of partial correlations between the observed neurons and inferred several latent units exerting linear effects on the observed neurons. We analyzed these networks of partial correlations and found the following: Whereas noise correlations were predominantly positive, the inferred partial correlations had a large fraction of negative values possibly reflecting inhibitory circuitry.  Moreover, we found that the inferred positive partial correlations exhibited stronger dependencies with respect to physical distances and differences in preferred orientations than noise correlations. In contrast, the inferred negative partial correlations were less selective.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
The covariance matrix is defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; the $p\times 1$ vector $x$ is a single observation the firing rates of $p$ neurons over time $\Delta t$; and $\mu$ is the vector of expected firing rates.  The usual way to estimate the covariance matrix is to calculate the \emph{sample covariance matrix} $C_{\sf 0}$ from the empirical sample of observations $x(t),\; t=1,\ldots,n$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom in one neuron's activity within the sample ($\nu=n-1$ if observations are independent). For noise correlations, the true mean $\mu$ and the sample mean $\bar x$ are conditioned on the stimulus. In this study, we estimate the true mean, $\mu$, using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than the sample correlation matrix, $C_{\sf 0}$.

@@

We considered four regularized estimators based on distinct families of reduced target estimates: $C_{\sf diag}$, $C_{\sf factor}$, $C_{\sf sparse}$, and $C_{\sf sparse+latent}$. In probabilistic models with exclusively linear dependencies, the targets estimates of these for estimators correspond to distinct families of graphical models (\figref{02}{\,Row 1}}).

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
        {\bf Row 1.} Graphical representations of the target estimates used of the four respective regularized covariance matrix estimators.  Recorded neurons are denotes as green spheres and latent units as light-shaded spheres.  The edges between them denote non-zero partial correlations.
        {\bf Row 1, A}.  In estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
        {\bf  Row 1, B.} In estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), favoring systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. In estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from linear associations between the observed units.
        {\bf  Row 1, D}.  In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to linear associations between the recorded cells \emph{and} with several latent units.
\\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices from row 2.  The structure of the sample correlation matrix is difficult to discern by eye.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with the matching structure converged to zero faster than the other estimators. Some exceptions occur when the sample is not large enough to reveal the true structure; in these cases estimators with simpler structures can be more efficient.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure of imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).  If sample correlations arise from spurious or contingent effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.

The target of estimator $C_{\sf factor}$ is the factor model $L + \Psi$, where $L$ is the $p\times p$ positive definite matrix with low rank and $\Psi$ is the diagonal matrix of independent variances.  The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (L + \Psi),
\end{equation}
has two hyperparameters: the rank of $L$ (\ie the number of latent factors) and the shrinkage intensity $\lambda$. The structure imposed by $C_{\sf factor}$ favors conditions in which the population activity is linearly driven by a number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{Row 1, B}).

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by a matrix whose inverse matrix is sparse: 
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.  The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).  We will sometimes refer to sparse partial interactions between neurons as simply \sq{interactions}.
